### 2018-01-12 cogsci/cogsci.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{The Role of Early Semantic and Phonological Networks in Noun Learning: a
Cross-linguistic Analysis}

\author{{\large \bf Abdellah Fourtassi} \\ \texttt{afourtas@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Yuan Bian} \\ \texttt{ybian.uiuc@gmail.com} \\ Department of Psychology \\ University of Illinois \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
Children tend to learn first the words that have a high density
neighborhood in the lexical network (i.e., words that are connected to a
variety of other words). However, the reason why connectivity influences
learning is still not very well understood. Previous studies suggested
this effect could be the consequence of the underlying learning
mechanism. However, these studies focused on semantic networks and
English data only, thus severely limiting the generality of the
findings. Here we provide a systematic analysis of network growth models
in both the phonological and the the semantic domains, and across 8
languages. We replicate the results obtained by Hills et al. (2009) with
English semantic networks, show that the same mechanism drives the
growth of both semantic and phonological networks, and demonstrate that
these effects hold cross-linguistically.

\textbf{Keywords:}
keywords: semantic network, phonological network, word learning,
semantic
\end{abstract}

\section{Introduction}\label{introduction}

A crucial question in language development is to understand why children
learn some words at an earlier age than others. Previous studies
documented various word properties that influence the age of
acquisition. For example, it was shown that, within a lexical category,
words that are more frequent in child-directed speech are acquired
earlier (Goodman, Dale, \& Li, 2008). Other predictors include word
length, the mean length of utterances in which the word occurs, and how
concrete the word is (see Braginsky, Yurovsky, Marchman, \& Frank,
2016). Besides these word-level properties, researchers found that the
lexical structure (that is, the network that specifies how words relate
to each other) also influences the learning process. For instance,
children tend to produce first the words that have higher neighborhood
density (i.g., high connectivity in the network) both at the
phonological and the semantic level (Stokes, 2010; Storkel, 2009).
However, it is still not very well understood \emph{why} connectivity in
the lexical structure influences learning. Steyvers \& Tenenbaum (2005)
suggested that this effect is nothing but the consequence of how these
structures get constructed in the child's mind. According to this
explanation (known as Preferential Attachment), highly connected words
in the child's lexicon tend to ``attract'' more words over time, in a
rich-get-richer scenario. Nonetheless, Hills, Maouene, Maouene, Sheya,
\& Smith (2009) found that what biases the learning is not the
connectivity in the child's internal lexicon but the connectivity in the
learning environment.

These analyses, however, focused on growth in semantic networks only. A
recent study by Stella, Beckage, \& Brede (2017) analysed both semantic
and phonological connectivity, but assumed only one growth model
(preferential acquisition). More importantly, almost all available
studies are limited in their generality because they focus almost
exlusively on English data. The novely of this study is that it
investigtes, for the first time, whether phonological networks grow by
preferential attachement or preferential acquisition. It also provides a
systematic comparision between network growth scenarios in both the
phonological and the semantic domains and assess their relative
contribution to the learning process. Moreover, it tests the generality
of the findings across eight languages.

The paper is organized as follows. First, we describe the dataset we
used, the procedure we followed to construct both semantic and
phonological networks, and the network growth scenarios we tested. Next,
we show the results of the various analyses we conducted comparing
network growth scenarios across languages. Then we compare the overall
contribution of the networks in the learning process relative to other
known predictors of age of acquisition (frequency and length). Finally
we discuss the major results in relation to known experimental facts in
language development.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/all_data-1}

}

\caption{\label{fig:data_all}Age of acquisition in the end-state network as predicted by the degree in this network. Results are shown in each language for the phonological network (top) and the semantic network (bottom). Each point is a word, with lines indicating linear model fits.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

\section{Networks}\label{networks}

\subsection{Data}\label{data}

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI). The latter consists of parent report vocabulary
checklists. We used the \emph{Words and Sentence} version which contains
the productive vocabulary of toddlers (age varied between 16 to 35
months). We included words that were produced by at least 50\% of
children, and we considered that these words were learned at the month
when this percentage was reached (Goodman et al., 2008). Moreover,
following previous studies (Hills et al., 2009; e.g., Storkel, 2009), we
restricted the analysis to the category of nouns, since there are known
discrepancies between the learning of nouns, predicates, and function
words (e.g., Braginsky et al., 2016). We obtained these nouns in eight
languages: Croatian, Danish, English, Italian, Norwegian, Russian,
Spanish, and Turkish. We used the subset of nouns that had entries in
the Florida Association Norms (see below). Since these norms are
available only in English, we mapped words onto translation equivalents
across CDI forms. This allowed us to use the English association norms
across languages. Table \ref{tab:stats} gives an overview of the data
used. The translation of non-English words is still an ongoing process.
Note, however, that all languages have at least 60\% of nouns
translated.

\begin{table}[H]
\centering
\begin{tabular}{rlrrr}
  \hline
 & language & total & translated & normed \\ 
  \hline
1 & Croatian & 253 & 177 & 170 \\ 
  2 & Danish & 295 & 198 & 187 \\ 
  3 & English & 296 & 296 & 274 \\ 
  4 & Italian & 311 & 203 & 194 \\ 
  5 & Norwegian & 305 & 193 & 186 \\ 
  6 & Russian & 311 & 311 & 285 \\ 
  7 & Spanish & 240 & 173 & 163 \\ 
  8 & Turkish & 293 & 175 & 164 \\ 
   \hline
\end{tabular}
\caption{\label{tab:stats}Total number of productive nouns in the CDI (left). We used a subset of these nouns that have available English translations (middle), as well as entries in the Free Association Norms (right).} 
\end{table}

With this data, we constructed both semantic and phonological networks.

\subsection{Semantic networks}\label{semantic-networks}

We used as index of semantic relatedness the Florida Free Association
Norms (Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected
by giving adult participants a word (the cue), and asking them to write
the first word that comes to mind (the target). For example, when given
the word ``ball'', they might answer with the word ``game''. A pair of
nodes were connected by a directed link from the cue to the target if
there was a cue-target relationship between these nodes in the
association norms. Each node was characterized by its ``indegree'',
which represents the number of links for which the word was the target.
Following Hills et al. (2009), we used as a proxy for the learning
environment, the network constructed using the full CDI data in each
language (e.g., in English this corresponds to the network by 30
months).

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/pred_ind_img-1}

}

\caption{\label{fig:pred_ind}Evaluation of individual network growth models using both  one-sample t-test over the z-score distribution of learned words (dotted), and model fitting (solid). Each dot represents, in the first case, the mean of the z-score distribution of the learned words with ranges representing 95\% confidence interval, and in the second case, the mean of the posterior distribution of the model's parameter, with ranges reppresenting 95\% credible intervals (computed using the highest density intervals).}\label{fig:pred_ind_img}
\end{figure*}
\end{CodeChunk}

\subsection{Phonological networks}\label{phonological-networks}

We used as a measure of phonological relatedness between two nodes the
Levenshtein (edit) distance between their phonological forms. The
measure counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another. We generated
approximate IPA forms from the orthographic transcription, across
languages, using the open source text-to-speech software
\textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}. In
previous studies, two nodes were linked if they had an edit distance of
1 (Stokes, 2010; Storkel, 2009). However, in these previous studies the
network was built using an adult vocabulary. Here, similar to the
approximation we made for the semantic network, we used the full set of
nouns in the CDI data as a proxy for phonological connectivity in the
learning environment. However, since the children's vocabulary contains
very few word pairs with an edit distance of 1, the resulting network
was too sparse and uninformative. Thus, we increased the threshold from
1 to 2, that is, two node were related if their edit distance was equal
to 1 or 2. Each node was characterized with its degree, i.e., the number
of links it shares with other words.

\subsection{Network growth}\label{network-growth}

We tested two network growth scenarios discussed in Hills et al. (2009).
The first one was \emph{Preferential Attachment} (PAT). According to
this mechanism, the network structure of known words predicts what words
will be learned next. A word is more likely to be learned if it is
linked to one of the highly connected nodes in this internal network.
The second mechanism was \emph{Preferential Acquisition} (PAC). In this
model, what predicts what word will enter the lexicon is not
connectivity in the internal network, but connectivity in the learning
environment. Figure \ref{fig:growth} shows an illustration of both
growth scenarios with the same simplified network. For PAT, each
candidate node was characterized with the average degree of the existing
nodes that it would attach to. Thus, according to PAT, the node N1 is
more likely to enter the lexicon first. For PAC, each candidate node was
characterized with its degree in the entire network. According th PAC,
the node N3 is more likely to enter the lexicon first.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/growth-1}

}

\caption{\label{fig:growth}Illustration of the two growth models with the same network. Filled circles (A to D) represent known words, and empty circles (N1 to N3) represent words that have not been learned yet. Back lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant (from Hills et al. 2009).}\label{fig:growth}
\end{figure}
\end{CodeChunk}

\section{Analysis}\label{analysis}

\subsection{Connectivity vs.~Age of
acquisition}\label{connectivity-vs.age-of-acquisition}

Preferential attachment and preferential acquisition provide different
generative scenarios for the same end-state. The latter is defined here
as the network corresponding to the oldest age for which we have CDI
data. However in both cases, it is predicted that nodes with the higher
degrees in this end-state are learned earlier than nodes with the lower
degrees. Figure \ref{fig:data_all} shows how the age of acquisition for
each word varies as a function of its degree (or indegree for the
semantic network) in the end-state network. In order to compare this
correlation across languages, the predictor (i.e., the degree) was
centered and scaled. The plot shows a negative correlation between the
month of acquisition and the degree (with the exception of the Russian
semantic network), indicating that nouns with higher degrees are
generally learned earlier.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/pred_comb_img-1}

}

\caption{\label{fig:pred_comb_img}Estimates of the growth models' parameters in each language. Dots represent the means of the posteriori distribution of the parameters, and ranges represent 95\% credible intervals (computed using the highest density intervals). }\label{fig:pred_comb_img}
\end{figure*}
\end{CodeChunk}

\subsection{Network growth models}\label{network-growth-models}

Following Hills et al. (2009), we evaluated the growth models in two
separate ways. The first evaluation consists in determining, in each
month, the model-dependent growth value distribution of all words that
could possibly be learned at this month, and then computing the z-score
of each learned word with respect to this distribution. For each model,
we tested if the distribution constituted by the z-scores of all learned
words was different from zero, using a one-sample t-test. The results
are shown in Figure \ref{fig:pred_ind}.

The second evaluation consists in fitting an explicit model to the data.
The purpose of this second evaluation is, first, to validate the
findings obtained with the one-sample t-test, and second, to determine
(in a next step) the relative contribution of each network growth
scenario (PAT vs.~PAC) across domains (phonological vs.~semantic).

We proceeded as follows. We calculated the probability that a word
\(w_i\), with a growth value \(d_i\) would enter the lexicon at a given
month, using a softmax function:

\begin{equation}
 p(w_i)= \frac{e^{\beta d_i}}{\sum_j e^{\beta d_j} }
\end{equation}

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher groth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. The normalization included words that were not yet
learned at the month the word at hand was acquired. We estimated the
parameter \(\beta\) using a Bayesian approach. We started with a uniform
distribution, and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been actually learned at that month (using formula
1). We ended up with a posterior distribution of \(\beta\), which we
summarized in Figure \ref{fig:pred_ind}.

The results from both evaluations were very similar and lead essentially
to the same conclusions. For the semantic networks, the results
replicate Hills et al.'s finding in English, which was that the semantic
network grows by PAC, not by PAT. Moreover, it generalizes this finding
to all other languages (with the exception of Russian). For the
phonological network, the results show that, generally speaking, PAC
fits the developmental trajectory better than PAT in all languages. We
note however that PAT, though weaker, fares better for the phonological
networks than it does for the semantic networks. In the former, it
predicts part of the growth process in some languages such as Croatian,
English, Norwegian and Russian. In the latter, it is rather universally
unpredictive.

Above we evaluated the network growth scenarios individually. As a next
step, we analysed their relative contribution to the learning process.
This was done through adding more fitting parameters to the model, that
is, by substituting \(\beta d_i\) in equation (1) with:
\[\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}\]
where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and
PhonoPAC. Using the same fitting technique, we obtained the values shown
in Figure \ref{fig:pred_comb_img}. In term of growth scenarios, we found
that PAC dominates the learning. In particular, though we found
previously that phonological PAT predicted part of the learning in some
languages when tested individually, here PAT appears to lose its
predictive power when pitted against phonological PAC. In terms of
domain (phonological vs.~semantic), both phonological and semantic
networks appear to contribute to learning, although the phonological
network appears to be stronger and more reliable across languages. In
summary, the findings show that both semantic and phonological networks
grow primarily through preferential acquisition, and that, generally
speaking, semantic and phonological networks both contribute to the
learning process.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/regressions_img-1}

}

\caption{\label{fig:regressions_img}Estimates of predictor coefficients by language. Values above 0 indicate a positive relationship (e.g. longer words tend to have a higher AoA), while values below 0 indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA). Ranges indicate 95\% confidence intervals.}\label{fig:regressions_img}
\end{figure*}
\end{CodeChunk}

\subsection{The big picture: comparison to other known predictors of age
of
acquisition}\label{the-big-picture-comparison-to-other-known-predictors-of-age-of-acquisition}

We saw that the way semantic and phonological information is structured
in the learning environment (i.e., PAC) contribute to noun learning
across languages. However, we know that other factors influence learning
as well (e.g., Braginsky et al., 2016). In what follows, we will
investigate how semantic and phonological connectivity interact with two
other factors. The first one is word frequency, a well studied factor
shown to predict the age of acquisition in a reliable fashion (e.g.
Goodman et al., 2008). The second factor is word length because it
correlates with phonological connectivity (Pisoni, Nusbaum, Luce, \&
Slowiaczek, 1985).

Since PAT was uninformative, we dropped it from this analysis. Thus, we
no longer needed to to fit the growth model month-by-month as in the
previous section. In fact, word utilities in the case of PAC are fixed,
they do not depend on previously learned words. A more direct way to
assess and compare the contribution of PAC in relation to other factors
is through conducting linear regressions, where connectivity in the
learning environment (at both the phonological and semantic level),
frequency and length predict the age of acquisition.

We used the frequency estimates from Braginsky et al. (2016) where
unigram counts were derived based on CHILDES corpora in each language.
For each word, counts included words that shared the same stem (e.g.,
cats counts as cat), or words that were synonymous (e.g.~father counts
as daddy). For word length, we used our generated IPA transcription.
This allowed us to get a more accurate estimate of word length across
languages than the number of characters in the orthographic
transcription.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/regressions_all_img-1}

}

\caption{\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined model. Ranges indicate 95\% confidence intervals. Fade color indicates estimates of PAC predictors in a model that does not include frequency and length as covariates.}\label{fig:regressions_all_img}
\end{figure}
\end{CodeChunk}

We conducted two analyses. On the one had, we fit a linear regression
for each language, and on the other hand, we fit a linear mixed-effect
to all the data pooled across languages, with language as a random
effect. Figure \ref{fig:regressions_img} shows the coefficient estimate
for each predictor in each language, and figure
\ref{fig:regressions_all_img} shows the coefficient estimates for all
languages combined (all predictors were centered and scaled). The
findings were as follows. Overall, frequency is the most reliable
predictor of age of acquisition. Word length predicts part of learning
in some languages such as Croatian and Norwegian, but not in others
(including English). It remains, however, a significant predictor in the
global model. As for the factors of interest, i.e., semantic and
phonological connectivities, we also found cross-linguistic differences.
The phonological connectivity contributes to learning in languages such
as Croatian, English and Russian, whereas semantic connectivity
contributes to learning in Turkish, Spanish and to some extent in
Croatian, but interestingly not in
English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous studiy, semantic connectivity was not tested in a model that included both frequency and length as covariates. Morever, our sample size is bigger.}.
Despite this cross-linguistic variation, both phonological and semantic
connectivity remain significant predictors in the global model.

\section{Discussion}\label{discussion}

The present study provided a comprehensive analysis of how lexical
connectivity influences the age of acquisition of nouns in toddlers. We
compared two network growth scenarios and assessed the relative
contribution of phonological and semantic information in 8 languages.
Part of the findings largely replicate the results obtained in Hills et
al. (2009), i.e., semantic networks (based on free associations) grow by
preferential acquisition, not by preferential attachment. Another
finding was that phonological networks also grow primarily by
preferential acquisition, especially when both scenarios (PAT and PAC)
were pitted against each other in the same model. These findings
generalize well across languages. Moreoever, both semantic and
phonological connectivity in the learning environment (i.e., PAC)
predict growth in a consistent way across many languages. However, when
pitted against other known predictors of age of acquisition (word
frequency and length), the effect of word connectivity shows a
cross-linguistic variation, predicting part of learning in some
languages, but not in others. Despite this coross-lingusitic variation,
both phonological and semantic connectivity contribute to the overall
learning (when data is pooled across languages).

A major results of the study is that children start by learning words
that have high phonological and semantic similarity with a variety of
other words in the learning environment, not in the available lexicon.
This suggests that children are sensitive to connectivity even without
having first acquired the connected words! How can children indirectly
detect highly connected words, and why would such words be more readily
learned? In the semantic case, free association can be predicted through
the patterns of word co-occurrence (Griffiths, Steyvers, \& Tenenbaum,
2007), meaning that highly connected words tend to be the words that
co-occur with many other words in various contexts. One possibility,
suggested by Hills, Maouene, Riordan, \& Smith (2010), is that the
referents of such words are more easily disambiguated as demonstrated,
for instance, by cross-situational learning experiments (Smith \& Yu,
2008). In the phonological case, connectivity is inherently correlated
with phonotactic probability (Vitevitch, Luce, Pisoni, \& Auer, 1999).
That is, highly connected words tend to be made of frequent sound
sequences, and vice versa. We know even infant (whose vocabulary is
still very rudimentary) develop a sensitivity for high frequency sound
sequences in the ambient native language (Jusczyk, Luce, \&
Charles-Luce, 1994). Interestingly, it was shown that phonotactic
probability facilitates learning and recognition of novel words in
toddlers and preschoolers (MacRoy-Higgins, Shafer, Schwartz, \& Marton,
2014; Storkel, 2001). Thus, children's ability to keep track of
co-occurrence statistics (both at the semantic and phonological level)
might explain their apparent sensitivity and preference for high
connectivity in the lexical network.

Finally, this study shares a number of limitations with previous studies
using similar datasets. For instance, we used normative age of
acquisition. However, individual children may not follow a unique
learning trajectory. The use of longitudinal data would allow us to
assess how the aggregate behavior relates to individual trajectories
(e.g., typical vs.~late talkers, Beckage, Smith, \& Hills, 2011).
Moreover, the results obtained in this study provide correlational but
not causal evidence. Thus, conclusion drawn from our network analysis
would require parallel evidence, especially through controlled
behavioral experiments.

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-beckage2011}{}
Beckage, N. M., Smith, L., \& Hills, T. T. (2011). Small worlds and
semantic network growth in typical and late talkers. \emph{PLOS ONE},
\emph{6}(5), 1--6.

\hypertarget{ref-braginsky2016}{}
Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2016).
From uh-oh to tomorrow: Predicting age of acquisition for early words
across languages. In \emph{Proceedings of the 38th Annual Conference of
the Cognitive Science Society}.

\hypertarget{ref-frank2017}{}
Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2017).
Wordbank: An open repository for developmental vocabulary data.
\emph{Journal of Child Language}, \emph{44}(3), 677--694.

\hypertarget{ref-goodman2008}{}
Goodman, J. C., Dale, P. S., \& Li, P. (2008). Does frequency count?
Parental input and the acquisition of vocabulary. \emph{Journal of Child
Language}, \emph{35}(3), 515--531.

\hypertarget{ref-griffiths07}{}
Griffiths, T. L., Steyvers, M., \& Tenenbaum, J. B. (2007). Topics in
semantic representation. \emph{Psychological Review}, \emph{114}(2),
2007.

\hypertarget{ref-hills2010}{}
Hills, T. T., Maouene, J., Riordan, B., \& Smith, L. B. (2010). The
associative structure of language: Contextual diversity in early word
learning. \emph{Journal of Memory and Language}, \emph{63}(3), 259--273.

\hypertarget{ref-hills2009}{}
Hills, T. T., Maouene, M., Maouene, J., Sheya, A., \& Smith, L. (2009).
Longitudinal analysis of early semantic networks: Preferential
attachment or preferential acquisition? \emph{Psychological Science},
\emph{20}(6), 729--739.

\hypertarget{ref-jusczyk1994}{}
Jusczyk, P. W., Luce, P. A., \& Charles-Luce, J. (1994). Infant's
sensitivity to phonotactic patterns in the native language.
\emph{Journal of Memory and Language}, \emph{33}(5), 630--645.

\hypertarget{ref-higgins2014}{}
MacRoy-Higgins, M., Shafer, V. L., Schwartz, R. G., \& Marton, K.
(2014). The influence of phonotactic probability on word recognition in
toddlers. \emph{Child Language Teaching and Therapy}, \emph{30}(1),
117--130.

\hypertarget{ref-nelson1998}{}
Nelson, D. L., McEvoy, C. L., \& Schreiber, T. A. (1998). The University
of South Florida word association, rhyme, and word fragment norms.
Retrieved from \url{http://w3.usf.edu/FreeAssociation/}

\hypertarget{ref-pisoni1985}{}
Pisoni, D. B., Nusbaum, H. C., Luce, P. A., \& Slowiaczek, L. M. (1985).
Speech perception, word recognition and the structure of the lexicon.
\emph{Speech Communication}, \emph{4}(1), 75--95.

\hypertarget{ref-smith2008}{}
Smith, L., \& Yu, C. (2008). Infants rapidly learn word-referent
mappings via cross-situational statistics. \emph{Cognition},
\emph{106}(3), 1558--1568.

\hypertarget{ref-stella2017}{}
Stella, M., Beckage, N. M., \& Brede, M. (2017). Multiplex lexical
networks reveal patterns in early word acquisition in children.
\emph{Scientific Reports}, \emph{7}.

\hypertarget{ref-steyvers2005}{}
Steyvers, M., \& Tenenbaum, J. B. (2005). The large-scale structure of
semantic networks: Statistical analyses and a model of semantic growth.
\emph{Cognitive Science}, \emph{29}(1), 41--78.

\hypertarget{ref-stokes2010}{}
Stokes, S. F. (2010). Neighborhood density and word frequency predict
vocabulary size in toddlers. \emph{Journal of Speech, Language, and
Hearing Research}, \emph{53}(3), 670--683.

\hypertarget{ref-storkel2001}{}
Storkel, H. L. (2001). Learning new words: Phonotactic probability in
language development. \emph{Journal of Speech, Language, and Hearing
Research}, \emph{44}(6), 1321--1337.

\hypertarget{ref-storkel2009}{}
Storkel, H. L. (2009). Developmental differences in the effects of
phonological, lexical and semantic variables on word learning by
infants. \emph{Journal of Child Language}, \emph{36}(2), 29--321.

\hypertarget{ref-vitevitch1999}{}
Vitevitch, M. S., Luce, P. A., Pisoni, D. B., \& Auer, E. T. (1999).
Phonotactics, neighborhood activation, and lexical access for spoken
words. \emph{Brain and Language}, \emph{68}(1), 306--311.

\end{document}