### 2018-11-12 journal/ms.tex
journal corrections

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word Learning as Network Growth: A Cross-linguistic Analysis},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Word Learning as Network Growth: A Cross-linguistic Analysis}

  \shorttitle{Word Learning as Network Growth}

  \author{Abdellah Fourtassi\textsuperscript{1}, Yuan Bian\textsuperscript{2}, \& Michael C. Frank\textsuperscript{1}}

  % \def\affdep{{"", "", ""}}%
  % \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University\\
          \textsuperscript{2} Department of Psychology, University of Illinois  }

  \authornote{
    Abdellah Fourtassi

    Department of Psychology

    Stanford University

    50 Serra Mall

    Jordan Hall, Building 420

    Stanford, CA 94301

    Correspondence concerning this article should be addressed to Abdellah
    Fourtassi, Postal address. E-mail:
    \href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}
  }

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along both the phonological and semantic
dimensions. Though this connectivity effect has been extensively
documented, little is known about the underlying developmental
mechanism. One view suggests that learning is primarily driven by a
network growth model where highly connected words in the child's early
lexicon attract similar words. Another view suggests that learning is
driven by highly connected words in the external learning environment,
instead of highly connected words in the early internal lexicon. The
present study tests both scenarios systematically in both the
phonological and semantic domains across 10 languages. We show that
external connectivity in the learning environment drives growth in both
production- and comprehension-based vocabularies, even controlling for
word frequency and length. This pattern of findings suggests a word
learning mechanism where children harness their statistical learning
abilities to (indirectly) detect and learn highly connected words in the
learning environment.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

  }

  \usepackage[sortcites=false,sorting=none]{biblatex}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

\section{Introduction}\label{introduction}

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from conducting controlled laboratory
experiments (e.g. Markman, 1990) to analyzing dense corpora capturing
language learning in context (e.g., B. C. Roy, Frank, DeCamp, Miller, \&
Roy, 2015). One strategy consists in documenting the timeline of words'
acquisition, and studying the properties that make words easy or hard to
learn. For example, within a lexical category, words that are more
frequent in child-directed speech are acquired earlier (J. C. Goodman,
Dale, \& Li, 2008). Other factors include word length, the mean length
of utterances in which the word occurs, and concreteness (see Braginsky,
Yurovsky, Marchman, \& Frank, 2016).

Besides these word-level properties, the structure of the lexicon (that
is, how words relate to each other) also predicts the age of acquisition
of words. The lexical structure can be characterized in terms of a
network where each node represents a word in the vocabulary, and each
link between two nodes represents a relationship between the
corresponding pair of words (e.g., Collins \& Loftus, 1975). Previous
studies have investigated early vocabulary structure by constructing
networks using a variety of word-word relations including shared
semantic features, target-cue relationships in free association norms,
co-occurrence in child directed speech, and phonological relatedness.
These studies have found that children tend to produce words that have
higher neighborhood density (i.e., high connectivity in the network)
earlier, both at the phonological and the semantic level (Carlson,
Sonderegger, \& Bane, 2014; Hills, Maouene, Riordan, \& Smith, 2010;
Hills, Maouene, Maouene, Sheya, \& Smith, 2009; Stella, Beckage, \&
Brede, 2017; Storkel, 2009).

While most studies have focused on the static properties of the lexical
network, a few have investigated the underlying developmental process.
In particular, Steyvers and Tenenbaum (2005) suggested that the observed
effects of connectivity are the consequence of how the lexical network
gets constructed in the child's mind. According to this explanation,
known as Preferential Attachment, highly connected words in the child's
lexicon tend to \enquote{attract} more words over time, in a
rich-get-richer scenario (Barabasi \& Albert, 1999). In other words,
what predicts word learning is the \emph{internal} connectivity in the
child's early lexicon. In contrast, Hills et al. (2009) suggested that
what biases the learning is not the connectivity in the child's internal
lexicon but, rather, \emph{external} connectivity in the learning
environment. They called this alternative explanation Preferential
Acquisition. For clarity of reading, we will call preferential
attachment the Internally-driven mechanism (INT), and preferential
acquisition the Externally-driven mechanism (EXT).
Figure~\ref{fig:growth} shows an illustration of both growth scenarios
with the same simplified network.

These two proposals represent two divergent ideas about the role of
lexical networks in acquisition. On the INT proposal, learning is driven
by known words with high connectivity to other known words (Figure 1,
left). Thus, the network structure is a causal factor in early word
learning, that is, children might be relying on the organization of
their past knowledge to determine future learning. For example, having a
rich and organized knowledge about the domain of dinosaurs would
facilitate the acquisition of new dinosaur-related words (Chi \& Koeske,
1983). In contrast, on the EXT approach, learning is driven by the
connectivity of words that are not known yet (Figure 1, right). Thus,
the relevant network structure is not internally represented by
children, and the related connectivity effect might be an epiphenomenon
of some low level properties of the linguistic input. For example,
highly connected words in the input could be more easily learned because
of their contextual diversity (e.g., Smith \& Yu, 2008) or because
caregivers emphasize such words in child-directed speech (MacWhinney,
2014).

\begin{figure}

{\centering \includegraphics[width=300px]{figs/growth3}

}

\caption{Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, the node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, the node 2 is more likely to enter the lexicon next.}\label{fig:growth}
\end{figure}

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesised by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is a profound finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 2) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured. In particular, words that are neighbors in the semantic
network (e.g., \enquote{cat}, \enquote{dog}) are not necessarily
neighbors in the phonological network, and vice versa. Does the
phonological network also predict word learning? Previous work did find
an effect of words' connectivity in the phonological network on their
age of learning (Carlson et al., 2014; Stella et al., 2017; Storkel,
2009). In other words, words learned earlier in life tend to sound
similar to many other words than a word learned later in life. However,
this finding is \emph{a priori} compatible with both INT and EXT, and
previous studies did not explicitly compare these two mechanisms. Here,
we investigate whether phonological networks, like semantic networks,
grow through EXT, or if they rather grow via INT
(Figure~\ref{fig:growth}).

Second, we study vocabularies measured using both comprehension and
production. Previous studies have found differences between these
vocabularies in terms of their content and rate of acquisition
(Benedict, 1979; Fenson et al., 1994). These differences may reflect the
fact that comprehension and production do not share the same
constraints. For instance, whereas comprehension depends on the ease
with which words are stored and accessed, production depends,
additionally, on the ease with which words are articulated, e.g.,
shorter words are produced earlier (Braginsky et al., 2016). By
investigating comprehension-based vocabularies, we assess the extent to
which the network growth mechanism captures general learning patterns
beyond the specific constraints of production.

Finally, we use developmental data in 10 languages. Lexical networks can
show more or less cross-linguistic variability along both the semantic
and phonological domains (Arbesman, Strogatz, \& Vitevitch, 2010; Youn
et al., 2016). Besides, cultures might differ in the way caregivers talk
to children (Cristia, Dupoux, Gurven, \& Stieglitz, 2017; Kuhl et al.,
1997), and this difference in the input could influence the children's
learning strategy. Thus, Cross-linguistic comparison is crucial to test
what mechanism is cognitively universal and is used by all children, and
what mechanism is specific to some patterns of learning that emerge due
to the particulars of a given language or culture (Bates \& MacWhinney,
1987; Slobin, 2014).

We test the growth scenarios using parent reports of children
vocabularies and their normative age of acquisition (Fenson et al.,
1994). Children may vary in their individual learning trajectories, but
the aggregate data leads to an average learning pattern which is highly
consistent. It is this normative trajectory that we model in the current
study, following the steps of previous research using similar datasets
(Braginsky et al., 2016; J. C. Goodman et al., 2008; Hills et al., 2010,
2009; Stella et al., 2017; Storkel, 2009)

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with AoA and shape of the distribution) and we explain how these
properties infom hypotheses about network growth. Next, we explicitly
fit the two hypothesized growth mechanisms to the data. We investigate
the extent to which the results obtained in Hills et al. (2009)
generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

\section{Networks}\label{networks}

\subsection{Data}\label{data}

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. Parent report is
a reliable and valid measure of children's vocabulary that allows for
the cost-effective collection of datasets large enough to test
network-based models of acquisition (Fenson et al., 1994). When filling
out a CDI form, caregivers are either invited to indicate whether their
child \enquote{understands} (comprehension) or \enquote{understands and
says} (production) each of about 400-700 words. For younger children
(e.g., 8 to 18 months in the English data), both comprehension and
production are queried, whereas for older children (16 to 36 months)
only production is queried. Thus, we use data from younger children to
test comprehension and data from older children to test production
across 10 languages. Following previous studies (Hills et al., 2009;
Storkel, 2009), we restrict our analysis to nouns. Table \ref{tab:stats}
gives an overview of the data.

\subsection{Age of acquisition}\label{age-of-acquisition}

For each word in the CDI data, we compute the proportion of children who
understand or produce the word at each month. Then we fit a logistic
curve to these proportions and determined when the curve crosses 0.5,
i.e., the age at which at least 50\% of children know the word. We take
this point in time to be each word's age of acquisition (Braginsky et
al., 2016; J. C. Goodman et al., 2008).

\begin{table}

\caption{\label{tab:stats}Statistics for dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Comprehension} & \multicolumn{2}{c}{Production} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
Language & Nouns & Ages & Nouns & Ages\\
\midrule
Croatian & 209 & 8-16 & 312 & 16-30\\
Danish & 200 & 8-20 & 316 & 16-36\\
English & 209 & 8-18 & 312 & 16-30\\
French & 197 & 8-16 & 307 & 16-30\\
Italian & 209 & 7-24 & 312 & 18-36\\
Norwegian & 193 & 8-20 & 316 & 16-36\\
Russian & 207 & 8-18 & 314 & 18-36\\
Spanish & 208 & 8-18 & 312 & 16-30\\
Swedish & 205 & 8-16 & 339 & 16-28\\
Turkish & 180 & 8-16 & 297 & 16-36\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Semantic networks}\label{semantic-networks}

We constructed semantic networks for English data following the
procedure outlined in Hills et al. (2009), as follows. We used as an
index of semantic relatedness the Florida Free Association Norms
(Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected by
giving adult participants a word (the cue), and asking them to write the
first word that comes to mind (the target). For example, when given the
word \enquote{ball}, they might answer with the word \enquote{game}. A
pair of nodes were connected by a directed link from the cue to the
target if there was a cue-target relationship between these nodes in the
association norms. The connectivity of a given node was characterized by
its \emph{indegree}: the number of links for which the word was the
target. To model growth from month to month, we constructed a different
network at each month, based on the words that have been acquired by
that month.

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across languages. Using
the same association data across languages does not necessarily mean
that the resulting networks will be the same across languages, or that
these networks will grow similarly. Indeed, though this approximation
assumes that the semantic similarity measure is universal----which is a
reasonable assumption (e.g., Youn et al., 2016), the set of words
acquired by children as well as the timeline of this acquisition can
still vary from language to language leading to possibility different
learning strategies.

\subsection{Phonological networks}\label{phonological-networks}

To construct phonological networks we first mapped the orthographic
transcription of words to their International Phonetic Alphabet (IPA)
transcriptions in each language, using the open source text-to-speech
software \textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}
We used the Levenshtein distance (also known as edit distance) as a
measure of phonological relatedness between two nodes. The measure
counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological networks
to word learning when these networks were built using a rich adult
vocabulary. However, since the focus of the current study is on the
mechanism of growth, the networks should be based on the children's
early vocabulary which, nevertheless, contains very few word pairs with
an edit distance of 1. Thus, we increased the threshold from 1 to 2,
that is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker that those done with a threshold of 2.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

\section{Analysis}\label{analysis}

\subsection{Static properties of the global
network}\label{static-properties-of-the-global-network}

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that are relevant to INT
and/or EXT. In order to compare various predictors on the same data, we
restrict the analysis to the subset of nouns for which we had both
semantic and phonological information in each language.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrAll-1} \caption{Age of production (A) and comprehension (B) in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.}\label{fig:corrAll}
\end{figure}

\subsubsection{Connectivity predicts the age of
acquisition}\label{connectivity-predicts-the-age-of-acquisition}

Connectivity in the global network is directly related to EXT as it
represents the explicit criterion this growth scenario uses to determine
what words should be learned first (Figure \ref{fig:growth}). Therefore,
a direct consequence of an EXT-like growth scenario is a correlation
between connectivity in the global network and the age of
acquisition.\footnote{This correlation is also compatible with INT, although the causality is reversed. Indeed, from the perspective of this growth scenario, higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time.}
Figure \ref{fig:corrAll} shows how the age of production and
comprehension for each word varies as a function of its degree (or
indegree for the semantic networks) as well as the correlation values.
For ease of visual comparison, the predictor (i.e., the degree) was
centered and scaled across languages.

The plots show, overall, a negative correlation between the month of
acquisition and the degree. In production data, the average correlation
across languages was -0.24 (SD=0.10) for the semantic networks and -0.30
(SD=0.05) for the phonological networks. In comprehension data, the
average correlation was -0.21 (SD=0.08) for the semantic networks and
-0.21 (SD=0.07) for the phonological networks. These results indicate
that nouns with higher degrees are generally learned earlier, thus
replicating previous findings in English (e.g., Storkel 2004, 2009;
Hills et al. 2009) and extending these findings to ten different
languages, generally, in both production- and comprehension-based
vocabularies.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/degreeDist-1} \caption{Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. The figure shows the results for both production and comprehension data. A perfect power-law distribution should appear as a straight line in this graph.}\label{fig:degreeDist}
\end{figure}

\subsubsection{Power-law degree
distribution}\label{power-law-degree-distribution}

We also analyzed the global network's degree distribution. The shape of
this distribution is particularly relevant to INT as this growth
scenario is known to generate networks with a power-law degree
distribution, i.e., a distribution of the form
\(p(k) \propto \frac{1}{k^{\alpha}}\) (Barabasi \& Albert, 1999). If the
network displays this property, this fact would suggest, but not prove,
an INT-like generative process. If the degree distribution does not
follow a power law, this would largely weaken the case for INT. The
log-log plots are shown in Figure \ref{fig:degreeDist}. We fit a power
law to each empirical degree distribution following the procedure
outlined in Clauset, Shalizi, and Newman (2009) and using a related R
package (poweRlaw, Gillespie, 2015).

\begin{table}

\caption{\label{tab:powerLawProd}Results of fitting a power law model to the degree (i.e., connectivity) distribution in each model for production data. Numbers indicate the cut-off degree, the scaling parameter alpha, and the p-value which quantifies the plausibility of the power law hypothesis. If the p-value is close to 1, a power law cannot be rejected as a plausible fit for the data.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{Phono.} & \multicolumn{3}{c}{Sem.} \\
\cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
Language & cut-off & alpha & p-value & cut-off & alpha & p-value\\
\midrule
Croatian & 4 & 2.18 & 0.123 & 4 & 2.55 & 0.881\\
Danish & 11 & 4.55 & 0.858 & 4 & 2.38 & 0.001\\
English & 20 & 9.14 & 0.511 & 5 & 2.66 & 0.132\\
French & 20 & 3.75 & 0.112 & 8 & 2.81 & 0.133\\
Italian & 9 & 9.45 & 0.780 & 4 & 2.93 & 0.608\\
Norwegian & 15 & 6.28 & 0.744 & 5 & 2.88 & 0.201\\
Russian & 8 & 4.20 & 0.541 & 24 & 5.61 & 0.723\\
Spanish & 13 & 8.75 & 0.736 & 4 & 2.98 & 0.460\\
Swedish & 11 & 4.68 & 0.103 & 4 & 2.49 & 0.171\\
Turkish & 8 & 3.26 & 0.375 & 4 & 2.87 & 0.925\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:powerLawComp}Results of fitting a power law model to the degree distribution in each model for comprehension data. Numbers indicate the cut-off degree, the scaling parameter alpha, and the p-value which quantifies the plausibility of the power law hypothesis. If the p-value is close to 1, a power law cannot be rejected as a plausible fit for the data.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{Phono.} & \multicolumn{3}{c}{Sem.} \\
\cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
Language & cut-off & alpha & p-value & cut-off & alpha & p-value\\
\midrule
Croatian & 2 & 2.06 & 0.020 & 5 & 2.67 & 0.895\\
Danish & 5 & 2.98 & 0.136 & 4 & 2.39 & 0.005\\
English & 13 & 5.16 & 0.235 & 4 & 2.64 & 0.765\\
French & 18 & 5.58 & 0.336 & 4 & 2.63 & 0.330\\
Italian & 8 & 10.27 & 0.909 & 4 & 2.88 & 0.688\\
Norwegian & 13 & 7.65 & 0.440 & 5 & 2.87 & 0.433\\
Russian & 5 & 3.97 & 0.854 & 8 & 3.91 & 0.952\\
Spanish & 5 & 3.01 & 0.085 & 5 & 3.11 & 0.552\\
Swedish & 9 & 6.75 & 0.102 & 5 & 2.81 & 0.713\\
Turkish & 9 & 5.73 & 0.958 & 4 & 3.13 & 0.887\\
\bottomrule
\end{tabular}
\end{table}

In brief, the analysis consisted in two steps. First, we derived the
optimal cut-off, \(k_{min}\), above which the distribution is more
likely to follow a power
law,\footnote{In natural phenomena, it is often the case that the power law applies only for values above a certain minimum.}
and we estimate the corresponding scaling parameter \(\alpha\). Second
we calculated the goodness-to-fit, which resulted in a \(p\)-value
quantifying the plausibility of the model. The results are shown in
Table \ref{tab:powerLawProd} for production data, and in Table
\ref{tab:powerLawComp} for comprehension data.

Overall, we could not reject the null hypothesis of a power-law
distribution: The \(p\)-value was generally above 0.1 in almost all
languages for both production and comprehension. That said, phonological
networks had relatively larger cut-offs than semantic networks. As was
suggested by Arbesman et al. (2010), these \enquote{truncated}
power-laws in phonological networks--as well as the observed
cross-linguistic variability in the value of the cut-offs--may reflect
the various constraints that exist on word formation such as the number
of phonemes in the language, the phonotactics (i.e., the way sound
sequences are arranged in words), and the length of words. Such
constraints may limit the number of words that are phonologically
similar, thus leading to distributions which decay faster than a
non-truncated power law.

In sum, the static properties of the global network are \emph{a priori}
compatible with both INT and EXT. In order to decide between these two
developmental scenarios, we need to fit explicit growth models to the
data.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/growthPred-1} \caption{Evaluation of growth scenarios (EXT: externally-driven, INT: internally-driven) for both semantic and phonological networks. Each point represents the mean of the posterior distribution of the growth parameter, with ranges representing 95\% credible intervals. Positive values mean that learning proceeds according to the predictions of the growth scenario, whereas negative values mean that learning proceeds in opposition to the predictions of the growth scenario.}\label{fig:growthPred}
\end{figure}

\subsection{Network growth models}\label{network-growth-models}

To test the network growth scenarios, we fit two growth models to the
data. We calculated the probability that a word \(w_i\), with a utility
value \(u_i\) would enter the lexicon at a given month, using a softmax
function:

\begin{equation}
 p(w_i)= \frac{e^{\beta u_i}}{\sum_j e^{\beta u_j} }
\end{equation}

\noindent where \(\beta\) is a fitted parameter that captures the
magnitude of the relationship between network parameters and growth
(analogous to a regression coefficient). A positive value of \(\beta\)
means that words with higher utility values \(u_i\) are acquired first,
and a negative value means that words with lower utility values are
acquired first (see Figure \ref{fig:growth} for an illustration of how
utilities values \(u_i\) are defined in each growth scenario). The
normalization includes all words that could be learned at that month.

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}.

First, the results replicate Hills et al.'s original finding regarding
the semantic network in English and production data, which is that this
network grows by EXT, not by INT. Second, our results show that,
generally speaking, this finding generalizes to comprehension, and holds
across languages. This generalization was obtained in both the
semantic\footnote{One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather it supports our initial approximation about the universality of the semantic similarity measure.}
and phonological domains.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/staticPred-1} \caption{Estimates of the relative contribution of each predictor of AoA in the regression model in each language. Results are shown for both production and comprehension data. Ranges indicate 95\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).}\label{fig:staticPred}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/staticAll-1} \caption{Estimates of the relative contribution of each predictor of AoA in the combined mixed-effects model with language as a random effect. Results are shown for both production and comprehension data. Ranges indicate 95\% confidence intervals. Dotted ranges indicate the estimates for the predictor in a separate model that includes only this predictor as a fixed effect.}\label{fig:staticAll}
\end{figure}

\subsection{Comparison to other predictors of age of
acquisition}\label{comparison-to-other-predictors-of-age-of-acquisition}

Above we showed that the way semantic and phonological information is
structured in the learning environment contributes to noun learning (via
EXT) across languages. However, we know that other factors influence
learning as well (e.g., Braginsky et al., 2016). Next we investigated
how semantic and phonological connectivity interact with two other
factors. The first one is word frequency, a well studied factor shown to
predict the age of acquisition in a reliable fashion (e.g. J. C. Goodman
et al., 2008). The second factor is word length, which was shown to
correlate with phonological connectivity: Shorter words are more likely
to have higher connectivity (Pisoni, Nusbaum, Luce, \& Slowiaczek, 1985;
Vitevitch \& Rodríguez, 2005).

Since we found INT to be uninformative, we dropped it from this
analysis, keeping only EXT. This simplified the model because we no
longer needed to fit growth month-by-month. The latter was a requirement
only for INT where the words' utilities varied from month to month,
depending on how connectivity changed in the growing internal network. A
more direct way to assess and compare the contribution of EXT in
relation to other word-level factors is through conducting linear
regressions, where connectivity in the learning environment, frequency
and length predict the age of acquisition.

For word length, we counted the number of phonemes in our generated IPA
transcription. For word frequency, we used the frequency estimates from
Braginsky et al. (2016) where unigram counts were derived based on
CHILDES corpora in each language (MacWhinney, 2014). For each word,
counts included words that shared the same stem (e.g., \enquote{cats}
counts as \enquote{cat}), or words that were synonymous (e.g.
\enquote{father} counts as \enquote{daddy}). Although these frequency
counts use transcripts from independent sets of children, they are based
on large samples, and this allows us to average out possible differences
between children and the specificities of their input (see J. C. Goodman
et al., 2008).

We conducted two analyses. We fit a linear regression for each language,
and we fit a linear mixed-effect model to all the data pooled across
languages, with language as a random effect. Figure \ref{fig:staticPred}
shows the coefficient estimate for each predictor in each language for
production and comprehension data. Figure \ref{fig:staticAll} shows the
coefficient estimates for all languages combined (all predictors were
centered and scaled).

The findings for the new predictors were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition in
both comprehension and production data and across languages, endorsing
results for nouns across a variety of analyses (Braginsky et al., 2016;
J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word length is more
predictive for production than comprehension (and this difference is
very clear in the global model), replicating previous work (Braginsky et
al., under review). Thus, word length seems to reflect the effects of
production's constraints rather than than comprehension's constraints,
i.e., longer words are harder to articulate but they may not be
significantly more difficult to store and access.

As for the factors of interest, i.e., semantic and phonological
connectivity, we found cross-linguistic differences. Connectivity
contributes to learning in some languages but not in other. In
particular, semantic connectivity does not explain variance in English
data beyond that explained by phonological connectivity, frequency and
length. This contrasts with the original finding in Hills et al. (2009).
However, this might be due to our using of a slightly different model
(which included word length as a covariate) and a larger dataset. That
said, and despite these apparent cross-linguistic differences, both
phonological and semantic connectivity are significant predictors in the
combined model.

\section{Discussion}\label{discussion}

This study provided an analysis of network growth during development. We
compared two network growth scenarios described in the pioneering work
of Steyvers and Tenenbaum (2005) and Hills et al. (2009). The first
scenario, INT (originally called Preferential Attachment), described a
rich-get-richer network growth model in which the current structure of
the learner's internal network determines future growth; the other, EXT
(originally called Preferential Acquisition) described a model in which
the external, global environmental network structure determines
learners' growth patterns. These two mechanisms represent two
fundamentally different accounts of lexical growth: One suggests that
future word knowledge is primarily shaped by the children's past
knowledge and its organization, whereas the other suggests that learning
is shaped, rather, by salient properties in the input regardless of how
past knowledge is organized. The present study tested the generality of
previous findings (Hills et al., 2010, 2009) by 1) investigating
phonological networks together with semantic networks, 2) testing both
comprehension- and production-based vocabularies, and 3) comparing the
results across 10 languages.

We found that the original findings reported in Hills et al. (2009)
generalize well across all these dimensions. First, just like semantic
networks, phonological networks grow via the externally-driven scenario
(EXT), not by the internally-driven mechanism (INT). Second,
comprehension-based vocabularies grow in a way similar to
production-based vocabularies. Finally, the findings were, overall,
similar across the 10 languages we tested. Although we find some
cross-linguistic variation when semantic and phonological networks were
pitted against frequency and length, this variability is to be taken
with a grain of salt as it might be exaggerated in our study by the
limited and partially-overlapping sample of nouns for each language. In
fact, both phonological and semantic connectivity are significant
predictors above and beyond frequency and length when data are pooled
across languages.

These findings corroborate the hypothesis that children start by
learning words that have high similarity to a variety of other words in
the learning environment, not in the child's available lexicon. This
hypothesis implies that children are sensitive to highly connected words
although they do not initially have access to the full network, thus
raising some important questions: What mechanism allows children to
distinguish highly connected words from other words? Besides, why would
highly connected words be easier to learn?

One possibility is that these patterns emerge from children's use of
statistical learning abilities (Aslin \& Newport, 2012; Saffran, Aslin,
\& Newport, 1996; Smith \& Yu, 2008). The term \enquote{statistical
learning} has been used in the developmental literature to describes the
process by which one acquires information about their environment
through keeping track of the frequency distribution of some elements
(e.g., words) in different contexts. An important property of this kind
of learning is that it occurs without explicit instructions and through
mere exposure to the input. Previous work in the line of research has
documented mechanisms which can explain the patterns found in the
current study.

For example, in the semantic domain, growth according to EXT can be
explained by a mechanism similar to cross-situational learning (Pinker,
2013; Smith \& Yu, 2008; Yurovsky \& Frank, 2015). According to this
mechanism, children track the co-occurrence of concrete nouns with their
possible semantic referents. The referent of a word heard in only one
naming situation can be ambiguous (e.g., when the word \enquote{ball} is
heard for the first time in the presence of both a ball and a chair),
but hearing the same word in a diversity of semantic contexts allows the
learner to narrow down the set of possible word-object mappings. In our
case, free association is related to contextual co-occurrence
(Griffiths, Steyvers, \& Tenenbaum, 2007), meaning that highly connected
words will tend to occur in a variety of speech and referential
contexts. This fact makes such words easier to learn because they have
more referential disambiguating cues across learning contexts, and
crucially, even without knowing the entire set of words with witch they
co-occur (hence the similarity with EXT). This possibility is supported
by the finding that words' diversity of occurrence in child directed
speech predicts their age of learning (Hills et al., 2010).

In the phonological case, network growth according to EXT is also
compatible with a scenario whereby children are tracking lwo level
statistical patterns, e.g., high probability sound sequences. Indeed,
connectivity in the phonological network is inherently correlated with
phonotactic probability (M. S. Vitevitch, Luce, Pisoni, \& Auer, 1999).
That is, highly connected words tend to be made of frequent sound
sequences. Children are sensitive to local phonotactic regularities
(Jusczyk, Luce, \& Charles-Luce, 1994) and this sensitivity might lead
them to learn higher-probability words more easily (Storkel, 2001). This
explanation is supported by computational simulations that shows how
learning general phonotactics patterns create ``well-worn paths'' which
allow the models to represent several distinct but phonologically
neighboring words (Dell, Juliano, \& Govindjee, 1993; Takac, Knott, \&
Stokes, 2017).

Besides using their own statistical learning skills, children could also
benefit from the way their caregivers speak. Perhaps the caregivers put
more emphasis on the words that are highly connected in \emph{their}
mature lexical network. This emphasis would guide children to learn
first these highly connected words even though children do not have
access to the distribution of words' connectivity in the final network.
Investigating this possibility would require further research on
caregiver-child interaction (MacWhinney, 2014; B. C. Roy et al., 2015),
examining what words are introduced over development and the extent to
which children's uptake is influenced by this input (Clark, 2007; Hoff
\& Naigles, 2002).

This work shares a number of limitations with previous studies using
similar research strategy and datasets. Chief among these limitations is
the fact that the age of word acquisition is computed using different
children at different ages (due to the fact that available CDI data is
mainly cross-sectional). Although this measure has proven highly
consistent (Fenson et al., 1994), it led us to focus on studying the
learning mechanism of the \enquote{average} child. Individual
trajectories, however, could show different learning patterns. For
example, using longitudinal data Beckage, Smith, and Hills (2011) found
differences between typical and late talkers in terms of the semantic
network structure. Besides, although our study endorses the
externally-driven account of network growth, this does not mean
individual children never use some variant of INT or some combination of
both INT and EXT (Beckage and Colunga, under review). For example, some
children develop \enquote{islands of expertise}, that is, well organized
knowledge about a certain topic (e.g., birds or dinosaurs). This prior
knowledge enables these children to learn new related words more easily
(e.g., Chi \& Koeske, 1983).

To conclude, our work validates and generalizes previous results in
early network development. It suggests that the advantage of highly
connected words may result, at least in the early stages of word
learning, from the operation of simpler mechanisms in both the semantic
and phonological domains. One question for future experimental work is
whether such correlational patterns of growth can be produced in
controlled behavioral experiments.

\vspace{1em}

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation.

\section{Disclosure statement}\label{disclosure-statement}

None of the authors have any financial interest or a conflict of
interest regarding this work and this submission.

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-arbesman2010}{}
Arbesman, S., Strogatz, S. H., \& Vitevitch, M. S. (2010). The structure
of phonological networks across multiple languages. \emph{International
Journal of Bifurcation and Chaos}, \emph{20}(03), 679--685.

\hypertarget{ref-aslin2012}{}
Aslin, R. N., \& Newport, E. L. (2012). Statistical learning: From
acquiring specific items to forming general rules. \emph{Current
Directions in Psychological Science}, \emph{21}(3).

\hypertarget{ref-barabasi99}{}
Barabasi, A.-L., \& Albert, R. (1999). Emergence of scaling in random
networks. \emph{Science}, \emph{286}(5439), 509--512.

\hypertarget{ref-bates1987}{}
Bates, E., \& MacWhinney, B. (1987). Competition, variation, and
language learning. \emph{Mechanisms of Language Acquisition}, 157--193.

\hypertarget{ref-beckage2011}{}
Beckage, N. M., Smith, L., \& Hills, T. T. (2011). Small worlds and
semantic network growth in typical and late talkers. \emph{PLOS ONE},
\emph{6}(5), 1--6.

\hypertarget{ref-benedict1979}{}
Benedict, H. (1979). Early lexical development: Comprehension and
production. \emph{Journal of Child Language}, \emph{6}(2), 183--200.

\hypertarget{ref-braginsky2016}{}
Braginsky, M., Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2016).
From uh-oh to tomorrow: Predicting age of acquisition for early words
across languages. In \emph{Proceedings of the 38th Annual Conference of
the Cognitive Science Society}.

\hypertarget{ref-carlson2014}{}
Carlson, M. T., Sonderegger, M., \& Bane, M. (2014). How children
explore the phonological network in child-directed speech: A survival
analysis of children?s first word productions. \emph{Journal of Memory
and Language}, \emph{75}, 159--180.

\hypertarget{ref-chi1983}{}
Chi, M. T., \& Koeske, R. D. (1983). Network representation of a child's
dinosaur knowledge. \emph{Developmental Psychology}, \emph{19}(1).

\hypertarget{ref-clark2007}{}
Clark, E. V. (2007). Young children's uptake of new words in
conversation. \emph{Language in Society}, \emph{36}(2).

\hypertarget{ref-clauset09}{}
Clauset, A., Shalizi, C. R., \& Newman, M. E. J. (2009). Power-law
distributions in empirical data. \emph{SIAM Review}, \emph{51}(4),
661--703.

\hypertarget{ref-collins1975}{}
Collins, A. M., \& Loftus, E. F. (1975). A spreading-activation theory
of semantic processing. \emph{Psychological Review}, \emph{82}(6).

\hypertarget{ref-cristia2017}{}
Cristia, A., Dupoux, E., Gurven, M., \& Stieglitz, J. (2017).
Child-directed speech is infrequent in a forager-farmer population: A
time allocation study. \emph{Child Development}.

\hypertarget{ref-dell1993}{}
Dell, G. S., Juliano, C., \& Govindjee, A. (1993). Structure and content
in language production: A theory of frame constraints in phonological
speech errors. \emph{Cognitive Science}, \emph{17}(2), 149--195.

\hypertarget{ref-fenson94}{}
Fenson, L., Dale, P. S., Reznick, J. S., Bates, E., Thal, D. J.,
Pethick, S. J., \ldots{} Stiles, J. (1994). Variability in early
communicative development. \emph{Monographs of the Society for Research
in Child Development}, \emph{59}(5).

\hypertarget{ref-frank2017}{}
Frank, M. C., Braginsky, M., Yurovsky, D., \& Marchman, V. A. (2017).
Wordbank: An open repository for developmental vocabulary data.
\emph{Journal of Child Language}, \emph{44}(3), 677--694.

\hypertarget{ref-gillespie15}{}
Gillespie, C. S. (2015). Fitting heavy tailed distributions: The
poweRlaw package. \emph{Journal of Statistical Software}, \emph{64}(2),
1--16. Retrieved from \url{http://www.jstatsoft.org/v64/i02/}

\hypertarget{ref-goodman2008}{}
Goodman, J. C., Dale, P. S., \& Li, P. (2008). Does frequency count?
Parental input and the acquisition of vocabulary. \emph{Journal of Child
Language}, \emph{35}(3), 515--531.

\hypertarget{ref-dippl}{}
Goodman, N., \& Stuhlmuller, A. (2014). The Design and Implementation of
Probabilistic Programming Languages. \url{http://dippl.org}.

\hypertarget{ref-griffiths07}{}
Griffiths, T. L., Steyvers, M., \& Tenenbaum, J. B. (2007). Topics in
semantic representation. \emph{Psychological Review}, \emph{114}(2),
2007.

\hypertarget{ref-hills2010}{}
Hills, T. T., Maouene, J., Riordan, B., \& Smith, L. B. (2010). The
associative structure of language: Contextual diversity in early word
learning. \emph{Journal of Memory and Language}, \emph{63}(3), 259--273.

\hypertarget{ref-hills2009}{}
Hills, T. T., Maouene, M., Maouene, J., Sheya, A., \& Smith, L. (2009).
Longitudinal analysis of early semantic networks: Preferential
attachment or preferential acquisition? \emph{Psychological Science},
\emph{20}(6), 729--739.

\hypertarget{ref-hoff2002}{}
Hoff, E., \& Naigles, L. (2002). How children use input to acquire a
lexicon. \emph{Child Development}, \emph{73}(2).

\hypertarget{ref-jusczyk1994}{}
Jusczyk, P. W., Luce, P. A., \& Charles-Luce, J. (1994). Infant's
sensitivity to phonotactic patterns in the native language.
\emph{Journal of Memory and Language}, \emph{33}(5), 630--645.

\hypertarget{ref-kuhl1997}{}
Kuhl, P. K., Andruski, J. E., Chistovich, I. A., Chistovich, L. A.,
Kozhevnikova, E. V., Ryskina, V. L., \ldots{} Lacerda, F. (1997).
Cross-language analysis of phonetic units in language addressed to
infants. \emph{Science}, \emph{277}(5326), 684--686.

\hypertarget{ref-macwhinney2014}{}
MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
talk, volume ii: The database}. Psychology Press.

\hypertarget{ref-markman90}{}
Markman, E. M. (1990). Constraints children place on word meanings.
\emph{Cognitive Science}, \emph{14}(1), 57--77.

\hypertarget{ref-nelson1998}{}
Nelson, D. L., McEvoy, C. L., \& Schreiber, T. A. (1998). \emph{The
University of South Florida word association, rhyme, and word fragment
norms}. Retrieved from \url{http://w3.usf.edu/FreeAssociation/}

\hypertarget{ref-pinker2013}{}
Pinker, S. (2013). \emph{Learnability and cognition: The acquisition of
argument structure}. MIT press.

\hypertarget{ref-pisoni1985}{}
Pisoni, D. B., Nusbaum, H. C., Luce, P. A., \& Slowiaczek, L. M. (1985).
Speech perception, word recognition and the structure of the lexicon.
\emph{Speech Communication}, \emph{4}(1), 75--95.

\hypertarget{ref-roy2015}{}
Roy, B. C., Frank, M. C., DeCamp, P., Miller, M., \& Roy, D. (2015).
Predicting the birth of a spoken word. \emph{Proceedings of the National
Academy of Sciences}, \emph{112}(41), 12663--12668.

\hypertarget{ref-saffran1996}{}
Saffran, J. R., Aslin, R. N., \& Newport, E. L. (1996). Statistical
learning by 8-month-old infants. \emph{Science}, \emph{274}(5294),
1926--1928.

\hypertarget{ref-slobin2014}{}
Slobin, D. I. (2014). \emph{The crosslinguistic study of language
acquisition} (Vol. 4). Psychology Press.

\hypertarget{ref-smith2008}{}
Smith, L., \& Yu, C. (2008). Infants rapidly learn word-referent
mappings via cross-situational statistics. \emph{Cognition},
\emph{106}(3).

\hypertarget{ref-stella2017}{}
Stella, M., Beckage, N. M., \& Brede, M. (2017). Multiplex lexical
networks reveal patterns in early word acquisition in children.
\emph{Scientific Reports}, \emph{7}.

\hypertarget{ref-steyvers2005}{}
Steyvers, M., \& Tenenbaum, J. B. (2005). The large-scale structure of
semantic networks: Statistical analyses and a model of semantic growth.
\emph{Cognitive Science}, \emph{29}(1), 41--78.

\hypertarget{ref-storkel2001}{}
Storkel, H. L. (2001). Learning new words: Phonotactic probability in
language development. \emph{Journal of Speech, Language, and Hearing
Research}, \emph{44}(6), 1321--1337.

\hypertarget{ref-storkel2009}{}
Storkel, H. L. (2009). Developmental differences in the effects of
phonological, lexical and semantic variables on word learning by
infants. \emph{Journal of Child Language}, \emph{36}(2), 29--321.

\hypertarget{ref-takac2017}{}
Takac, M., Knott, A., \& Stokes, S. (2017). What can neighbourhood
density effects tell us about word learning? Insights from a
connectionist model of vocabulary development. \emph{Journal of Child
Language}, \emph{44}(2).

\hypertarget{ref-vitevitch2005}{}
Vitevitch, M. S., \& Rodríguez, E. (2005). Neighborhood density effects
in spoken word recognition in spanish. \emph{Journal of Multilingual
Communication Disorders}, \emph{3}(1).

\hypertarget{ref-vitevitch1999}{}
Vitevitch, M. S., Luce, P. A., Pisoni, D. B., \& Auer, E. T. (1999).
Phonotactics, neighborhood activation, and lexical access for spoken
words. \emph{Brain and Language}, \emph{68}(1), 306--311.

\hypertarget{ref-youn2016}{}
Youn, H., Sutton, L., Smith, E., Moore, C., Wilkins, J. F., Maddieson,
I., \ldots{} Bhattacharya, T. (2016). On the universal structure of
human lexical semantics. \emph{Proceedings of the National Academy of
Sciences}, \emph{113}(7), 1766--1771.

\hypertarget{ref-yurovsky2015}{}
Yurovsky, D., \& Frank, M. C. (2015). An integrative account of
constraints on cross-situational learning. \emph{Cognition}, \emph{145}.

\end{document}