### 2019-09-13 main_analyses.Rmd

Libraries
```{r}
  library(broom)
  library(purrr)
  library(readr)
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(wordbankr)
  library(stringr)
  library(feather)
  library(lme4)
  library(boot)
  library(langcog)
  library(ggthemes)
  library(nlme)
  library(rwebppl)
  library(jsonlite)
  library(Hmisc)
  library(poweRlaw)
  library(HDInterval)
  library(kableExtra)
```

Import helper functions and probabilstic models

```{r}
#Load helper functions
source(paste(getwd(),"/helper_functions/all_helper.r",sep = ""), chdir = T)

#Load probabilsitic models
source(paste(getwd(),"/models/all_models.r",sep = ""), chdir = T)

```

Load data and pre-process
```{r}

languages = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish")

admins <- get_administration_data() %>%
  select(data_id, age, language, form) %>%
  filter(language %in% languages)

items <- get_item_data() %>%
  filter(type == "word", lexical_class == "nouns") %>%
  filter(language %in% languages)

items_by_inst <- split(items, paste(items$language, items$form, sep = "_"))

get_inst_data <- function(inst_items) {
  inst_lang <- unique(inst_items$language)
  inst_form <- unique(inst_items$form)
  inst_admins <- filter(admins, language == inst_lang, form == inst_form)
  get_instrument_data(language = inst_lang ,
                      form = inst_form,
                      administrations = inst_admins,
                      items = inst_items$item_id,
                      iteminfo = inst_items
                      ) %>%
    filter(!is.na(age)) %>%
    mutate(produces = !is.na(value) & value == "produces",
           understands = !is.na(value) & (value == "understands" | value == "produces")) %>%
    select(-value) %>%
    gather(measure, value, produces, understands) %>%
    filter((measure == "understands" & form == "WG") | (measure == "produces" & form == "WS") ) %>%
    mutate(language = inst_lang,
           form = inst_form)

}

data_raw <- map(items_by_inst, get_inst_data)

data_all <- bind_rows(data_raw) %>%
  rename(item = num_item_id) %>%
  group_by(language, form, measure,
             lexical_category, lexical_class, uni_lemma,  item, definition, 
             age) %>%
  summarise(num_true = sum(value, na.rm = TRUE),
              num_false = n() - num_true,
              prop = mean(value, na.rm = TRUE))

#feather::write_feather(data_all, "saved_data/data_all.feather")

#I am taking out measure to keep only produce
data_all <- feather::read_feather("saved_data/all_data.feather")

```

Here compute the statistics 
language, measure, number of nouns, age limits, and number  
```{r}

#Number of children

#Number of children
n_child <- data_all %>%
  mutate(sum_child =  num_true+num_false) %>%
  group_by(language, measure, item) %>%
  summarise(children = sum(sum_child)) %>%
  distinct(language, measure, children)

mysummary <- data_all %>%
  group_by(language, measure) %>%
  summarise(min_age = min(age),
            max_age = max(age),
            nouns = n_distinct(item)) %>%
  left_join(n_child) %>%
  mutate(age_range = paste(min_age, max_age, sep = "-")) %>%
  select(-min_age, -max_age)  %>%
  group_by(language, measure) %>%
  nest() %>%
  spread(measure, data) %>%
  unnest(.sep = "_") %>%
  select(language, un = understands_nouns, ua=understands_age_range, uc= understands_children,
                   pn = produces_nouns, pa=produces_age_range, pc=produces_children)

feather::write_feather(mysummary, "saved_data/stats_new.feather")

kable(mysummary, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for data from Wordbank.",
      col.names = c("Language",  "Nouns", "Ages", "Nouns", "Ages")) %>%
  add_header_above(c("", "Comprehension" = 2, "Production" = 2)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

```

Compute Age of Acqusition (AoA) of words

```{r}

fit_inst_measure_uni <- function(inst_measure_uni_data) {
  ages <- min(inst_measure_uni_data$age):max(inst_measure_uni_data$age)
  model <- glm(cbind(num_true, num_false) ~ age, family = "binomial",
               data = inst_measure_uni_data)
  fit <- predict(model, newdata = data.frame(age = ages), se.fit = TRUE)
  aoa <- -model$coefficients[["(Intercept)"]] / model$coefficients[["age"]]
  constants <- inst_measure_uni_data %>%
    ungroup()%>%
    select(language, form, measure, lexical_category, lexical_class, uni_lemma, item, definition) %>%
    distinct()

  props <- inst_measure_uni_data %>%
    ungroup() %>%
    select(age, prop)

  data.frame(age = ages,
             fit_prop = inv.logit(fit$fit),
             fit_se = fit$se.fit,
             aoa = aoa, language = constants$language,
             form = constants$form,
             measure = constants$measure,
             uni_lemma = constants$uni_lemma,
             item = constants$item,
             definition = constants$definition) %>%
    left_join(props)
}

list_by_item <- data_all %>%
  # make this filtering at the beginning, not here
  split(paste(.$language, .$form, .$measure, .$item))

data_aoa <- map(list_by_item, fit_inst_measure_uni) %>%
  bind_rows()

feather::write_feather(data_aoa, "saved_data/data_aoa.feather")

aoa <- feather::read_feather("saved_data/data_aoa.feather")  %>%
  select(language, form, measure, uni_lemma, item, definition, aoa) %>%
  distinct()

#I am taking out measure to keep only produce
#data_aoa2 <- feather::read_feather("saved_data/aoa_data.feather") %>%
#  rename(definition = defintion) %>%
#  select(language, form, measure, uni_lemma, item, definition, aoa) %>%
#  distinct()

```

Compute the growth data structue (based on AoA data)

```{r}

#make a growth dataframe
make_data_growth <- function(word_aoa) {

  measure <- unique(word_aoa$measure)
  lang <- unique(word_aoa$language)

  ages<- (word_aoa %>%
    arrange(age) %>%
    group_by(age) %>%
    summarise(n = n()))$age

  df <- data.frame()
  for (i in ages) {
    rem_words <- word_aoa %>% filter(age >= i)
    rem_lemma <- c(rem_words$uni_lemma)
    rem_def <- c(rem_words$definition)
    rem_item<- c(rem_words$item)
    corr_age <- rep(i, times = length(rem_lemma))
    curr_df <- data.frame(corr_age, rem_item, rem_lemma, rem_def)
    df <- rbind(df, curr_df)
  }  
  df <- df %>% rename(uni_lemma = rem_lemma, definition=rem_def, item=rem_item)%>%
    left_join(word_aoa %>% select(item, age)) %>%
    mutate(learned = as.numeric(age == corr_age)) %>%
    select(corr_age, item, definition, uni_lemma,learned) %>%
    rename(age = corr_age) %>%
    arrange(age, item) %>%
    mutate(language = lang,
           measure = measure)
  return(df)
}

aoa_by_measure_lang <- data_aoa %>%
  mutate(age = round(aoa)) %>%
  select(-form, -aoa) %>%
  split(paste(.$language, .$measure))

data_growth <- map(aoa_by_measure_lang, make_data_growth) %>%
  bind_rows()

#Save at this point
#feather::write_feather(data_growth, "saved_data/data_growth.feather")
data_growth <- feather::read_feather("saved_data/data_growth.feather")

```

Construct the netwroks and derive the dynamic predictors: PAC and PAT
```{r}

#Here I chose to use the full data for the phono netwotk and filter later for the unilemmas only when doing what? the combined regression?, thus the utility will be more precise for the resulting subset

words_growth <- data_growth %>%
    #filter(!is.na(uni_lemma)) %>% #only keep unilemmas that intersect with free association data
  filter(age > 5)

growth_by_meas <- words_growth %>%
  split(paste(.$measure))

growth_by_meas_lang <- words_growth %>%
  split(paste(.$measure, .$language))

#Semantic netwrok 
##################

sem_net_fun  <- function (growth_meas_lang){

  first_age<- growth_meas_lang$age[1]

  lemma_list<- growth_meas_lang %>%
      trim_all_unilemma() %>% #For semantic netowtks only the unilmma
      filter(age==first_age) %>%
      select(item, uni_lemma)

assoc_pairs<- make_assoc_pairs(lemma_list = lemma_list)

assoc_PAC<- PAC_generator(vocab_age = growth_meas_lang, word_pairs = assoc_pairs) %>% 
      rename(PAC_assoc = value) %>% select(-definition, -uni_lemma)

 assoc_PAT<- PAT_generator(vocab_age = growth_meas_lang, word_pairs = assoc_pairs) %>% 
      rename(PAT_assoc = value) %>% select(-definition, -uni_lemma)

 sem_growth <- growth_meas_lang %>%
   left_join(assoc_PAC) %>%
   left_join(assoc_PAT)

 return(sem_growth)

}

sem_red <-  map(growth_by_meas_lang, sem_net_fun) %>%
  bind_rows()

#Save 
#feather::write_feather(sem_red, "saved_data/sem_red.feather")
sem_red <- feather::read_feather("saved_data/sem_red.feather")

#Phonological Networks
######################

phono_net_fun <- function (growth_meas_lang) {

  lang <- unique(growth_meas_lang$language)
  #meas <- unique(growth_meas_lang$measure)

  first_age<- growth_meas_lang$age[1]

  def_list<- growth_meas_lang %>%
      trim_all_definition() %>% 
      filter(age==first_age) %>%
      select(item, definition)

  phono_pairs<- make_IPA_pairs(def_list = def_list, lang = lang) %>%
    mutate(language = lang) %>%

  #Threshold the phonetic distance (we take t=2, becuase t=1 is too sparse) 
  threshold <- 1

  phono_PAC <- PAC_generator(vocab_age = growth_meas_lang, word_pairs = phono_pairs %>% IPA_threshold(threshold)) %>% 
      rename(PAC_phono_t2 = value) %>% select(-definition, -uni_lemma)

  phono_PAT <- PAT_generator(vocab_age = growth_meas_lang, word_pairs = phono_pairs %>% IPA_threshold(threshold)) %>% 
      rename(PAT_phono_t2 = value) %>% select(-definition, -uni_lemma)

 # phono_nt_PAC <- PAC_generator(vocab_age = growth_meas_lang, word_pairs = phono_pairs %>% IPA_continuous()) #%>% 
 #    rename(PAC_phono = value) %>% select(-definition, -uni_lemma)

  phono_growth <- growth_meas_lang %>%
   left_join(phono_nt_PAC) 
   left_join(phono_PAC) %>%
   left_join(phono_PAT)

  return(phono_growth)
}

phono_red_nt <-  map(growth_by_meas_lang, phono_net_fun) %>%
  bind_rows()

feather::write_feather(phono_red_nt, "saved_data/phono_red_nt.feather")

feather::write_feather(phono_red, "saved_data/phono_red_edit1.feather")
phono_red <- feather::read_feather("saved_data/phono_red_nt.feather")

#Combine data 
data_growth_net <- words_growth %>%
  left_join(sem_red) %>%
  left_join(phono_red)

feather::write_feather(data_growth_net, "saved_data/data_growth_net_nt.feather")

```

Here do Mantel tests
```{r}

phono_net_fun <- function (growth_meas_lang) {

  lang <- unique(growth_meas_lang$language)
  #meas <- unique(growth_meas_lang$measure)

  first_age<- growth_meas_lang$age[1]

  def_list<- growth_meas_lang %>%
    filter(!is.na(uni_lemma)) %>%
    trim_all_definition() %>% 
    filter(age==first_age) %>%
    select(item, definition)

  item_unilemma <- growth_meas_lang %>%
    filter(!is.na(uni_lemma)) %>%
    select(item, uni_lemma)

  phono_pairs<- make_IPA_pairs(def_list = def_list, lang = lang) %>%
    mutate(language = lang) %>%
    left_join(item_unilemma) %>%
    rename(item0 = item,
           item = pair,
           item.uni_lemma = uni_lemma) %>%
    left_join(item_unilemma) %>%
    rename(pair.uni_lemma = uni_lemma,
           pair = item,
           item = item0)

  return(phono_pairs)
}

#Ici mantel test
phono_pairs_all <- map(growth_by_meas_lang, phono_net_fun) %>%
  bind_rows()

feather::write_feather(phono_pairs_all, "saved_data/phono_pairs_all.feather")

phono_pairs_dist <- phono_pairs_all %>%
  mutate(link1 = as.numeric(dist<=1)) %>%
  mutate(link2 = as.numeric(dist<=2)) %>%
  mutate(link_cont = ifelse(dist==0, 1, 1/dist))

#Pairwise comparision: each time, take the overlap and do the test

data_prod <- data_aoa %>%
  filter(measure =='produces')

data_by_lang <- split(phono_pairs_dist$language)

pair_data <- function(lang1){

  pair_cor_data <- function (lang2){

  lang1_unilemmas <- unique(lang1$uni_lemma)
  lang2_unilemmas <- unique(lang2$uni_lemma)
  uni_intersect <-  intersect(lang1_unilemmas, lang2_unilemmas)

  data_lang1 <- lang1 %>%
    filter(!is.na(uni_lemma),
           uni_lemma %in% uni_intersect) %>%
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa1=mean(aoa))

  data_lang2 <- lang2 %>%
    filter(!is.na(uni_lemma),
           uni_lemma %in% uni_intersect) %>%
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa2=mean(aoa))

  both <- data_lang1 %>%
    left_join(data_lang2)

  mycorr <- cor.test(rank(both$aoa1), rank(both$aoa2), method = 'kendall')
  val=mycorr$estimate
  p=mycorr$p.value
  language1 = unique(lang1$language) 
  language2 = unique(lang2$language)

  data.frame(language1=language1, language2=language2, tau=val, sig = p)

}
  map(data_by_lang, pair_cor_data) %>%
    bind_rows()

}

```

Compute other static predictors (frequency and length) besides PAC

```{r}

data_static <- data_growth_net %>%
  distinct(measure, language, uni_lemma, definition, item, PAC_assoc, PAC_phono) %>%
  rename(sem_deg = PAC_assoc, 
         phono_deg =  PAC_phono)

#Word length

words_len <- data_static %>%
  trim_all_definition() %>%
  rowwise() %>%
  mutate(IPA=Speak(lang = language, word = definition)) %>%
  trim_IPA_completely() %>%
  mutate(length = str_count(IPA)) %>%
  select(-definition, -uni_lemma, -sem_deg, -phono_deg)

#Word frequency: we use frequency values derived from Cross-lingusitic corpora in CHILDES (from Mika e al. 2016) 
load("saved_data/uni_joined.RData")

frequency_mika <- uni_joined %>%
  filter(lexical_classes =='nouns') %>%
  distinct(language, uni_lemma, frequency) %>%
  mutate(lang_temp = ifelse(language == "French (Quebec)", "French (Quebecois)", language)) %>%
  select(-language) %>%
  rename(language = lang_temp)

words_freq <- data_static %>%
  left_join(frequency_mika) %>%
  select(-definition, -uni_lemma, -sem_deg, -phono_deg)

aoa_items <- aoa %>%
  select(language, measure, item, aoa) # Here I should round the AoA?

#Combine predictors

data_static_pred  <- data_static %>%
  left_join(words_len) %>%
  left_join(words_freq) %>%
  left_join(aoa_items) #%>%
  #select(-definition, -uni_lemma)

#Use feather here
################

#Combine with full proportion-based data  (for the second regression which fit the entire production curve)

data_static_pred$language <- plyr::mapvalues(data_static_pred$language, 
                                 from = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", "Russian", "Spanish", "Swedish", "Turkish"))

data_static_prop <- data_all %>%
  left_join(data_static_pred)

```

## Correlations

Correlatinon degree/AoA using unilemmas

```{r}

#data_all$language <- plyr::mapvalues(data_all$language, 
#                                 from = c("Croatian","Danish","English (American)", "French (Quebecois)", #"Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
#                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", #"Russian", "Spanish", "Swedish", "Turkish"))

#Sem_deg is defined over unilemmas,  phono_deg is defined over definitions, and AoA is defined over items 
#For a given defintion we can have more than one item (polysemy), for one unilemma we can have more than one definiton (if the semantic distinction is not very important), or we might have 0 unilemma (if the meaning of the defintion is not part of the main cross-lingusitc lemmas)
#The following analyses predict the items' AoA. I keep duplicated values of the predictors for different items (when they have the same defintions and/or unilemma) since I consider that this is the actual (albeit braod) information brough by the precitors

#Restrict analysis to unilemmas (i.e., definitons which have translations in English). This allows us to comapre various predictors on the same set of data
unilemmas <- data_static_pred %>%
  select(-IPA) %>%
  filter(!is.na(uni_lemma)) %>% 
  filter(!is.na(sem_deg)) %>% #only keep unilemmas that intersect with free association data 
  filter(aoa > 5) #These AoA are  artifact of the glm fitting (likely due to very few data point, I need to check these)

unilemmas_nt <-unilemmas %>%
  filter(phono_deg != 'Inf')

#Impute missing values and center/scale
uni_scale <- unilemmas_nt %>%
  group_by(measure, language) %>%
  mutate_at(c('sem_deg', 'phono_deg', 'length', 'frequency'), funs(as.numeric(Hmisc::impute(.)))) %>%
  mutate_at(c('sem_deg', 'phono_deg', 'length', 'frequency'), funs(as.numeric(scale(.))))

feather::write_feather(uni_scale, "saved_data/data_static_scaled_nt.feather")
#phono_red <- feather::read_feather("saved_data/phono_red.feather")

data_long <- uni_scale %>%
  gather(predictor, value, sem_deg:frequency) %>%
  filter(predictor == "sem_deg" | predictor == "phono_deg")

correlations <- data_long %>%
  group_by(measure, language, predictor) %>%
  summarise(cor = round(cor(aoa, value), 2))

#Reproduce the first analyses of correlation, (degree distribution later), then regressions both language specific and aggragate, 
#Then go to the intercation based regression, and aggregate, show that there is no significant interaction with age. And also show

```

Here do some analysis
```{r}
#Regression with phono (edit=1)

#Regression with phono (1/edit)
```

Production

```{r fig.width=10, fig.height=5}
plot_correlation_prod <- ggplot(subset(data_long, measure == "produces"), aes(x=value, y=aoa))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  #coord_cartesian(xlim=c(-1,10))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces"), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("AoA")

plot_correlation_prod

```

Comprehension
```{r fig.width=10, fig.height=5}
plot_correlation_comp <- ggplot(subset(data_long, measure == "understands"), aes(x=value, y=aoa))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands"), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("AoA")

plot_correlation_comp

```

##Static netwrok properties

### Degree distribution
Import the analysese from cogsci paper

```{r}

#function that takes data growth and does power law analyses

#Data for plot
degreeDist <- data.frame(matrix(ncol = 5, nrow = 0))
dist_names <- c("x", "y", "dimension", "language", "measure")
colnames(degreeDist) <- dist_names

#Parameters and test
degreeTest <- data.frame(matrix(ncol = 6, nrow = 0))
test_names <- c("xMin", "alpha", "pVal", "dimension","language", "measure")
colnames(degreeTest) <- test_names

powerLaw_fun <- function(data_growth_meas_lang, analysis){

  lang <- unique(data_growth_meas_lang$language)
  meas <- unique(data_growth_meas_lang$measure)

  data_meas_lang <- data_growth_meas_lang %>%
    select(measure, language, sem_deg, phono_deg) %>%
    dplyr::rename(Sem = sem_deg, Phono = phono_deg) %>%
    dplyr::filter (!is.na(Sem),!is.na(Phono))

  ##Semantic network

  #fit and derive parameters for power law
  semList <- data_meas_lang$Sem[data_meas_lang$Sem != 0]
  sem_pl = displ$new(semList)
  sem_est = estimate_xmin(sem_pl)
  sem_pl$setXmin(sem_est)
  sem_dist = plot(sem_pl) %>%
    mutate(dimension='Sem', language  =  lang, measure = meas)

  #bootstrap to get p-value
  sem_boot = bootstrap_p(sem_pl, no_of_sims=1000, threads=2)

  sem_test <- data.frame(as.numeric(sem_pl$xmin), as.numeric(sem_pl$pars), as.numeric(sem_boot$p), 'Sem', lang, meas)
  colnames(sem_test) <- test_names

  ##Phonological network

  #fit and derive parameters for power law
  phonoList <- data_meas_lang$Phono[data_meas_lang$Phono != 0]
  phono_pl = displ$new(phonoList)
  phono_est = estimate_xmin(phono_pl)
  phono_pl$setXmin(phono_est)
  phono_dist = plot(phono_pl) %>%
    mutate(dimension='Phono', language = lang, measure = meas)

  #bootstrap to get p-value
  phono_boot = bootstrap_p(phono_pl, no_of_sims=1000, threads=2)

  phono_test <- data.frame(as.numeric(phono_pl$xmin), as.numeric(phono_pl$pars), as.numeric(phono_boot$p), 'Phono', lang, meas)
  colnames(phono_test) <- test_names

  dist_meas_lang <- bind_rows(sem_dist, phono_dist)
  test_meas_lang <- bind_rows(sem_test, phono_test)

  #return(dist_meas_lang)

  if (analysis == 'distribution') {

    return(dist_meas_lang)

  } else if (analysis == 'test') {

    return(test_meas_lang)

  } else {

    print("Please specify the analysis type ('distribution' or 'test')")

  }

}

#Split by measure and language

data_by_meas_lang <- unilemmas %>%
  split(paste(.$measure, .$language))

degree_dist  <- map2(data_by_meas_lang, 'distribution', powerLaw_fun) %>%
  bind_rows()

feather::write_feather(degree_dist, "saved_data/degree_dist.feather")

degree_test  <- map2(data_by_meas_lang, 'test' , powerLaw_fun) %>%
  bind_rows()

feather::write_feather(degreeDist, "saved_data/degree_dist.feather")
degree_dist <- feather::read_feather("saved_data/degree_dist.feather")

feather::write_feather(degree_test, "saved_data/degree_test.feather")
degree_test <- feather::read_feather("saved_data/degree_test.feather")

#plot cumulative distributions
ggplot(data = degree_dist,  aes(x=x, y=y, col=dimension))+
  facet_grid(measure ~ language)+#, scales = "free") +
  geom_point(#aes(colour = lexical_class),
    #colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme(aspect.ratio = 1)
  #scale_x_continuous(limits=c(-2,5))+
  #geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  #scale_colour_solarized(name = "")

```

## Growth mechanisms

PreProcessing growth data strcutire for network growth models

```{r}

#data_growth_net <- feather::read_feather("saved_data/data_growth_net.feather")
data_growth_net_as_ws <- feather::read_feather("../asNet_local/saved_data/data_growth_net_AS_WS.feather")

#Filtering 
data_growth_pred <- data_growth_net %>%
  filter(!is.na(uni_lemma)) %>% #use only words with english translation 
  filter(!is.na(PAC_assoc)) %>% # keep unilemma which intersect with free association data
  filter(!is.na(PAC_phono_t2)) # In case a word is not transcribed phonologically (there should not be any)

  # replace NA in PAT with O (NA is when the utility cannot be computed, e.g., a candidate word is not linked to any previousely learned word, as opposed to being linked to a previously learned word with a degree=0)

data_growth_pred$PAT_assoc[is.na(data_growth_pred$PAT_assoc)] <- 0
data_growth_pred$PAT_phono_t2[is.na(data_growth_pred$PAT_phono_t2)] <- 0

#Scale predictors

data_growth_pred$language <- plyr::mapvalues(data_growth_pred$language, 
                                 from = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", "Russian", "Spanish", "Swedish", "Turkish"))

data_growth_pred <- data_growth_pred %>%
  group_by(measure, language) %>%
  mutate_at(c('PAC_assoc', 'PAT_assoc', 'PAC_phono_t2', 'PAT_phono_t2'), funs(as.numeric(scale(.)))) %>%
  select(-definition) #removing defintions because they cause JSon parser to fail in rWebppl (especially the (') in French)

#data_growth_pred_as <- feather::read_feather("../asNet_local/saved_data/data_growth_pred_wg_as_test")

```

Here we train the probabilistic models

```{r}
#Define the models

#Load probabilsitic models
source(paste(getwd(),"/models/all_models.r",sep = ""), chdir = T)

#Isolateed predictors
model_semPAT <-  paste(helper, sem_PAT, optimize, sep = '\n')
model_semPAC <-  paste(helper, sem_PAC, optimize, sep = '\n')
model_phonoPAT <-  paste(helper, phono_PAT, optimize, sep = '\n')
model_phonoPAC <-  paste(helper, phono_PAC, optimize, sep = '\n')

#Combined predictors
model_semPAC_phonoPAC_semPAT_phonoPAT <-  paste(helper, semPAC_phonoPAC_semPAT_phonoPAT, optimize, sep = '\n')

#Make a list of theses models
model_list = list(model_semPAT, model_semPAC, model_phonoPAT, model_phonoPAC,
                  model_semPAC_phonoPAC_semPAT_phonoPAT
                  )

model_name=c('semPAT', 'semPAC', 'phonoPAT', 'phonoPAC',
                                        'model_semPAC_phonoPAC_semPAT_phonoPAT')

```

Compute posteriors
```{r}

#Function that computes the posterior

posterior_fun <- function(data_growth_meas_lang) {

  lang <- unique(data_growth_meas_lang$language)
  meas <- unique(data_growth_meas_lang$measure)

  posteriors_lang <- vector("list", length = length(model_list))

   #for (i in 1:1) {
   for (i in 1:length(model_list)) {

    posterior_model <- NULL
    print(model_name[[i]]) 
    print(meas)
    print(lang)

    while (typeof(posterior_model)=="NULL") {
      posterior_model <- webppl(model_list[[i]],
                                data = data_growth_meas_lang,
                                data_var = "data",
                                inference_opts = list(
                                  method = "MCMC",
                                  samples = 500, 
                                  burn = 500),
                                model_var = "AOA",
                                output_format = "webppl")
    }

    posteriors_lang[[i]] <- posterior_model %>%
      mutate(measure = meas,
             language = lang,
             model_name = model_name[i])
  }

  return(bind_rows(posteriors_lang))

}

#Split by measure and language

data_growth_by_meas_lang <- data_growth_pred %>%
  split(paste(.$measure, .$language))

posterior <- map(data_growth_by_meas_lang, posterior_fun) %>%
  bind_rows()

feather::write_feather(posterior, "saved_data/posterior_as_ws.feather")
#posterior <- feather::read_feather("saved_data/posterior.feather")

#data_growth_meas_lang <- data_growth_by_meas_lang[[1]]

#posterior_fun(data_test)

```

Extact coefficients from the posterior

```{r}

param <- posterior %>%
  select(starts_with("value.parameters"), measure, language, model_name) %>%
  gather(alpha, value, value.parameters.alpha1:value.parameters.alpha6) %>%
  mutate(alpha = gsub("value.parameters.", "", alpha)) %>%
  dplyr::filter(alpha != "alpha5",
         alpha != "alpha6") %>%
  rename(model = model_name)

#feather::write_feather(param, "saved_data/param.feather")
#param <- feather::read_feather("saved_data/param.feather")

param_summary <- param %>%
  group_by(measure, language, model, alpha) %>%
  summarise(mean = mean(value),
            median = median(value),
            quantile_lower = quantile(value, c(0.025, 0.975))["2.5%"],
            quantile_upper = quantile(value, c(0.025, 0.975))["97.5%"],
            hdi_lower = hdi(value)['lower'],
            hdi_upper = hdi(value)['upper']
  )

param_summary_ind <- param_summary %>%
  dplyr::filter(model != "model_semPAC_phonoPAC_semPAT_phonoPAT",
         alpha == "alpha1") %>%
  mutate(Test = 'Individual') %>%
  select(-alpha)

param_summary_com <- param_summary %>%
  ungroup() %>%
  dplyr::filter(model == "model_semPAC_phonoPAC_semPAT_phonoPAT") %>%
  select(-model) %>%
  rename(model = alpha) %>%
  mutate(Test = 'Combined')

param_summary_com$model <- plyr::mapvalues(param_summary_com$model, 
                                 from = c("alpha1", "alpha2","alpha3","alpha4"), 
                                 to = c("semPAC", "phonoPAC", "semPAT", "phonoPAT"))

param_summary_all <- param_summary_ind %>%
  bind_rows(param_summary_com)

#feather::write_feather(param_summary_all, "saved_data/growth_preds.feather")
#param_summary_all <- feather::read_feather("saved_data/growth_preds.feather")

```

Plot growth data (production)

```{r fig.width=5, fig.height=2}
ggplot(subset(param_summary_all, measure == "produces" & Test =="Individual"), aes(x = model , y = mean)) +
  geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.3)))+
  scale_colour_solarized() +
  theme_bw()+
  theme(aspect.ratio = 0.7)

```

Plot growth data (comprehension)

```{r fig.width=5, fig.height=2}
ggplot(subset(param_summary_all, measure == "understands" & Test =="Individual"
              ), aes(x = model , y = mean)) +
  geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model, linetype=Test),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.3)))+
  scale_colour_solarized() +
  theme_bw()+
  theme(aspect.ratio = 0.7)

```

## Compare PAC to other predictors of AoA

```{r}

sem_formula <- as.formula("aoa ~ sem_deg")
phono_formula <- as.formula("aoa ~ phono_deg")
all_formula <- as.formula("aoa ~ frequency + length + sem_deg + phono_deg")

formulas <- list(sem_formula,  phono_formula, all_formula)

coefs_all <- data.frame()

for (formula in formulas){

#Model
model_reg <- function(data) {
   lm(formula, data = data)
}

#CI
ci_reg <- function(data) {
   confint(lm(formula, data = data))
 }

uni_scale <- feather::read_feather("saved_data/data_static_scaled_nt.feather")
#uni_scale_edit <- feather::read_feather("saved_data/data_static_scaled_edit1.feather")

reg <- uni_scale %>%
  group_by(measure, language) %>%
  nest() %>%
  mutate(model = map(data, model_reg)) %>%
  mutate(conf = map(data, ci_reg))

coefs_form <- reg %>%
  mutate(coefs = map(model, broom::tidy)) %>%
  mutate(ci = map(conf, broom::tidy)) %>%
  select(measure, language, coefs, ci) %>%
  unnest() %>%
  select(-.rownames) %>%
  dplyr::filter(term != "(Intercept)") %>%
  rename(predictor = term) %>%
  mutate(Test = ifelse(toString(formula[3])=='frequency + length + sem_deg + phono_deg', 'Combined', 'Individual'))

 coefs_all <- bind_rows(coefs_all, coefs_form)

}

feather::write_feather(coefs_all, "saved_data/coefs_all_nt.feather")

```

Plot for production
```{r}
plot_reg_prod <- ggplot(subset(coefs_all, measure =="produces"), aes(x = predictor, y = estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., y = estimate, col = predictor, linetype=Test), 
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  facet_wrap(~language, ncol=4)  +
  coord_flip() +
  guides(colour=FALSE)+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7)

plot_reg_prod

```

Plot for comprehension
```{r}
plot_reg_comp <- ggplot(subset(coefs_all, measure =="understands"), aes(x = predictor, y = estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., y = estimate, col = predictor, linetype=Test),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+

  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  facet_wrap(~language, ncol=4)  +
  coord_flip() +
  guides(colour=FALSE)+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7)

plot_reg_comp

```

Regression for data aggregated across all languages

```{r}

#Regression with all predictors and language as a random effect
#reg_agg <- lme(aoa ~ frequency + length + sem_deg + phono_deg , random=~1|language,  data= uni_scale)

sem_formula <- as.formula("aoa ~ sem_deg + (1 | language)")
phono_formula <- as.formula("aoa ~ phono_deg + (1 | language)")
all_formula <- as.formula("aoa ~ frequency + length + sem_deg + phono_deg +  (1 | language)")

formulas <- list(sem_formula,  phono_formula, all_formula)

coefs_agg_all <- data.frame()

for (formula in formulas){

reg_agg_prod <- lmer(formula,  data= subset(uni_scale, measure == "produces"))
reg_agg_comp <- lmer(formula,  data= subset(uni_scale, measure == "understands"))

#Extract coeficnet and confidence intervals
coef_agg_fun <- function (model, meas_name) {
conf <- confint(model, method ="Wald") 
conf <- data.frame(predictor = row.names(conf), conf) %>%
  dplyr::filter(!(predictor %in% c('.sig01', '.sigma', '(Intercept)')))

coef <- coef(summary(model))
data.frame(predictor = row.names(coef), coef) %>%
  dplyr::filter(predictor !='(Intercept)') %>%
  left_join(conf) %>%
  mutate(measure = meas_name)
}

coef_agg_prod <- coef_agg_fun(reg_agg_prod, 'production')
coef_agg_comp <- coef_agg_fun(reg_agg_comp, 'comprehension')

coef_agg_form <- coef_agg_prod %>%
  bind_rows(coef_agg_comp) %>%
  mutate(Test = ifelse(toString(formula[3])=='frequency + length + sem_deg + head(phono_deg + (1 | language)', 'Combined', 'Individual'))

 coefs_agg_all <- bind_rows(coefs_agg_all, coef_agg_form)

}

feather::write_feather(coefs_agg_all, "saved_data/static_preds_all_nt.feather")
coefs_agg_all <- feather::read_feather("saved_data/static_preds_all_nt.feather")

plot_agg <- ggplot(coefs_agg_all, aes(x = predictor, y = Estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., y = Estimate, col = predictor, linetype=Test),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  facet_wrap(~measure)  +
  coord_flip() +
  guides(colour=FALSE)+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7)

plot_agg

```
# Supplementary material

## Interaction with age

To examine interaction with age, we need to predict curve of acquisition instead of just AoA 
```{r}
#Start with reproducing Mika's analysis

predictors <- c("sem_deg", "phono_deg", "length", "frequency")
#predictors <- c("frequency", "length")
###Do we want to separate by 'form'? let's omit 'form' for now and focus on priduction (the case of understading is more problematic)

cols <- c("language", "measure", "uni_lemma", "lexical_class", "lexical_class", "defintion", "age",
          "num_true", "num_false", "prop", predictors)

load("saved_data/uni_joined.RData")
uni_joined$language <- plyr::mapvalues(uni_joined$language, 
                                 from = c("Croatian","Danish","English (American)", "French (Quebec)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", "Russian", "Spanish", "Swedish", "Turkish"))

uni_joined <- uni_joined %>%
  dplyr::filter(lexical_classes =='nouns') %>%
  rename(length = num_phons)

#Impute and scale (it's also good to have a sense of the missing data)
#uni_model_data <- uni_joined %>%
#  dplyr::filter(lexical_classes =='nouns') %>%
uni_model_data <- data_static_prop %>%
  #filter(language =="English (American)") %>%
  dplyr::filter(!is.na(uni_lemma)) %>% 
  #filter(!is.na(sem_deg)) %>% #only keep unilemmas that intersect with free association data 
  dplyr::filter(aoa > 5) %>%
  #group_by(language, measure) %>%
  group_by(measure, language) %>%
  mutate_at(vars(!!predictors), funs(as.numeric(Hmisc::impute(.)))) %>% #To fill NA 
  mutate(unscaled_age = age) %>%
  mutate_at(vars(c(!!predictors, "age")), funs(as.numeric(scale(.)))) %>%# Center and scale predictors 
  ungroup()

```

Interaction with age
```{r}
#This works for each language separately,

lang_model_fun <- function(lang, lang_data) {
   print(sprintf("Fitting glmer for %s...", lang))
   interaction_formula <- as.formula(
     #sprintf("prop ~ (age | item) +  %s",
     sprintf("cbind(num_true, num_false) ~ (age | uni_lemma) +  %s",
             paste(sprintf("age * %s", predictors), collapse = " + "))
   )

   glmer(interaction_formula, family = "binomial", data = lang_data, 
         control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e7))
         )

}

lang_models <- uni_model_data %>%
  group_by(measure, language) %>%
  nest() %>%
  mutate(model = map2(language, data, lang_model_fun))# %>%
  #mutate(conf = map2(language, data, lang_coef_fun))

all_model_fun <- function(data) {
   interaction_formula <- as.formula(
     sprintf("cbind(num_true, num_false) ~ (age | item) + (sem_deg + phono_deg + length + frequency | language) + %s",
             paste(sprintf("age * %s", predictors), collapse = " + "))
   )

   glmer(interaction_formula, family = "binomial", data = data, 
         control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e7))
         )

}

all_model <- uni_model_data %>%
  group_by(measure) %>%
  nest() %>%
  mutate(model = map(data, all_model_fun))#

all_coefs_net <- all_model %>%
  mutate(coefs = map(model, broom::tidy)) %>%
  #mutate(ci = map(conf, broom::tidy)) %>%
  select(measure, coefs) %>%
  unnest() %>%
  dplyr::filter(term != "(Intercept)", term != "age", group == 'fixed') %>%
  mutate(interaction = ifelse(grepl(":", term), "interaction with age",
                                "main effect"),
           term = gsub("age:", "", term)) %>%
  rename(predictor = term) %>%
  mutate(ci_low = estimate-1.96*std.error,
         ci_up = estimate+1.96*std.error)

lang_coefs_net <- lang_models %>%
  mutate(coefs = map(model, broom::tidy)) %>%
  #mutate(ci = map(conf, broom::tidy)) %>%
  select(measure, language, coefs) %>%
  unnest() %>%
  dplyr::filter(term != "(Intercept)", term != "age", group == 'fixed') %>%
  mutate(interaction = ifelse(grepl(":", term), "interaction with age",
                                "main effect"),
           term = gsub("age:", "", term)) %>%
  rename(predictor = term) %>%
  mutate(ci_low = estimate-1.96*std.error,
         ci_up = estimate+1.96*std.error)

mean_coefs <- lang_coefs_net %>%
  group_by(predictor, measure, interaction) %>%
  summarise(mean = mean(estimate))

```

```{r}
ggplot(lang_coefs_net, aes(x = estimate, y = predictor)) +
  #facet_grid(measure ~ interaction,  scales="free_x") +
  facet_grid(measure ~ interaction) +
  geom_point(aes(colour = predictor), size = 1, alpha = 0.4) +
  #geom_point(aes(x = mean, colour = predictor), size = 3, data = mean_coefs) +
  geom_point(aes(x = estimate, colour = predictor), size = 3, data = all_coefs_net) +
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed") +
  scale_fill_solarized(guide = FALSE) +
  scale_colour_manual(guide = FALSE,
                      values = rev(solarized_palette(num_coefs))) +
  ylab("") +
  scale_x_continuous(name = "Coefficient estimate")
```

```{r fig.width=5, fig.height=2}
ggplot(data=subset(lang_coefs_net, measure== 'produces'), aes(x = predictor, y = estimate)) +
  facet_grid(~ interaction, scales="free_x") +
  geom_pointrange(aes(ymin = ci_low, ymax = ci_up, col=language), 
                  position = position_dodge(width = .5),
                  size=0.3,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip()

ggplot(data=subset(lang_coefs_net, measure== 'understands'), aes(x = predictor, y = estimate)) +
  facet_grid(~ interaction, scales="free_x") +
  geom_pointrange(aes(ymin = ci_low, ymax = ci_up, col=language), 
                  position = position_dodge(width = .5),
                  size=0.3,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip()

```